Inference for Trans-dimensional Bayesian Models with Diffusive Nested Sampling

### 2014-12-05 Documents/Paper/ms.tex
New plot
references

\section{Sinusoidal example}\label{sec:sinewaves}
In this section we demonstrate a seemingly simple example which exhibits
a phase transition, making a standard MCMC approach difficult. Suppose
a signal is composed of $N$ sinusoids, of different periods, amplitudes,
and phases. The signal is observed at various times $\{t_i\}_{i=1}^N$ with
noise, and we want to use the resulting data to infer the number of sinusoids
$N$, along with the periods $\{T_i\}_{i=1}^N$, amplitudes $\{A_i\}_{i=1}^N$,
and phases $\{\phi_i\}_{i=1}^N$.
This kind of model has many applications, and has been solved analytically
under certain sets of assumptions \citep[see e.g.][]{bretthorst}. Similar
models have been used in many different fields
\citep[e.g.][]{2003AIPC..659....3B, 2005PhRvD..72b2001U, 2007ApJ...654..551B,
2009MNRAS.395.2226B}.
Note that this problem is also very closely related to the
problem of detecting exoplanet signals in radial velocity data, which has
attracted a lot of research attention in recent years
\citep[e.g.][]{2007ApJ...654..551B, gregory, fengji, 2011MNRAS.415.3462F}.

@@

@@

>>>>

\section{Sinusoidal example}\label{sec:sinewaves}
In this section we demonstrate a seemingly simple example which exhibits
a phase transition, making a standard MCMC approach difficult. Suppose
a signal is composed of $N$ sinusoids, of different periods, amplitudes,
and phases. The signal is observed at various times $\{t_i\}_{i=1}^N$ with
noise, and we want to use the resulting data to infer the number of sinusoids
$N$, along with the periods $\{T_i\}_{i=1}^N$, amplitudes $\{A_i\}_{i=1}^N$,
and phases $\{\phi_i\}_{i=1}^N$.
This kind of model has many applications, and has been solved analytically
under certain sets of assumptions \citep[see e.g.][]{bretthorst, 2014arXiv1412.0467M}.
Similar
models have been used in many different fields
\citep[e.g.][]{2003AIPC..659....3B, 2005PhRvD..72b2001U, 2007ApJ...654..551B,
2009MNRAS.395.2226B}.
Note that this problem is also very closely related to the
problem of detecting exoplanet signals in radial velocity data, which has
attracted a lot of research attention in recent years
\citep[e.g.][]{gregory, fengji, 2011MNRAS.415.3462F}.

@@

\begin{figure}
\begin{center}
\includegraphics[scale=0.5]{trace_logl.pdf}
\caption{A ``trace plot'' of the log likelihood in the SineWave model, for an
MCMC chain designed to sample the posterior distribution. The two phases
have log-likelihoods around -730 and -745 respectively, however transitions
between the two phases are quite uncommon.
\label{fig:trace_logl}}
\end{center}
\end{figure}

@@

\bibitem[Mortier et al.(2014)]{2014arXiv1412.0467M} Mortier, A., Faria, 
J.~P., Correia, C.~M., Santerne, A., Santos, N.~C.\ 2014.\ BGLS: A Bayesian 
formalism for the generalised Lomb-Scargle periodogram.\ ArXiv e-prints 
arXiv:1412.0467
